---
title: "Demonstrate the Computation of Indirect Multiplers and Impacts of construction spending from a Colorado EEIO model"
output:
  md_document:
    variant: gfm
---

# Demonstrate the Computation of Type I Multipliers and Related Impacts of construction spending from a Colorado EEIO model

This script builds a Colorado USEEIO State model for 2020 with extensions for Jobs, Greenhouse Gas Emissions, and Value Added components.

It requires _useeior_ >= v1.5.1 which will be installed from github using the remotes library if _useeior_ doesn't exist already in your R environment. The kableExtra package is used for formatting the tables for output here.

Your R environment needs to be setup to use the Project directory to knit this file (usually from RStudio).

```{r setup}
# Load a local R file with an install_useeior() function
source("R/utils.R")
install_useeior() # This install
library(useeior)
library(kableExtra)
```

Then we build a custom state EEIO model for Colorado using _useeior_ and a [new custom model specification file](model_specs/COEEIOv1.0-s-JGV-20.yml).
All the raw files including the relevant StateIO models and National totals of Employment and [GHGs by sector](https://doi.org/10.23719/1529805), and Indicators (for GHGs to convert to kg CO2e) tables are retrieved from the EPA Data Commons and stored locally in your machine-defined `appdata` directory under _stateio_ and _flowsa_, and _lciaformatter_, respectively (each are EPA Tools for Industrial Ecology modeling).

After building the model, it will store a version in RDS and Excel formats in an output directory. 
If you've already built it once, it will load the stored RDS version.

```{r buildCOmodel}
dir <- file.path("output")
local_model_RDS <- "COEEIOv1.0-s-JGV-20.rds"
dir_local_model_RDS <- file.path(dir,local_model_RDS)
if (!file.exists(dir_local_model_RDS)) {
  CO <- buildModel("COEEIOv1.0-s-JGV-20","model_specs/COEEIOv1.0-s-JGV-20.yml")  
  if (!dir.exists(dir)) dir.create(dir)
  saveRDS(CO,dir_local_model_RDS)
  writeModeltoXLSX(CO,outputfolder=dir)
} else {
  CO <- readRDS(dir_local_model_RDS)
}
```

Assume a scenario of $2 million in spending in construction in Colorado.

Get the Type I (indirect) output multiplier, defined here has the sum of total requirements table column for that sector and region of interest.

```{r outputmultiplers}
# Create a custom demand vector for CO construction
sector <- c("23/US-CO")
demand <- c(2E6)
regions <- c("CO","RoUS","All US")

output_multiplier_CO <- sum(CO$L_d[grep("US-CO",row.names(CO$L_d)),sector])
output_multiplier_RoUS <- sum(CO$L_d[grep("RoUS",row.names(CO$L_d)),sector])
output_multiplier <- output_multiplier_CO + output_multiplier_RoUS

kable(data.frame(Output_Multiplier=c(output_multiplier_CO,output_multiplier_RoUS,output_multiplier), row.names=regions))
```

Look at total output stimulated in CO, the rest of the U.S., and all the U.S.
```{r}
total_output_stimulated_CO <- demand*output_multiplier_CO
total_output_stimulated_RoUS <- demand*output_multiplier_RoUS
total_output_stimulated <- demand*output_multiplier

kable(data.frame(`Direct_Indirect_Output_Stimulated_$`=c(total_output_stimulated_CO,total_output_stimulated_RoUS,total_output_stimulated), row.names=regions))
```

Now calculate the direct and indirect impacts on greenhouse gas emissions, jobs supported, and value added.

```{r computeresults}
y <- setNames(demand, sector)
y <- formatDemandVector(y, CO$L)
# Use a direct perspective calculation which will show the impact in the sectors in which it occurs
result <- calculateEEIOModel(CO, perspective = "DIRECT", use_domestic_requirements=TRUE, demand = y)
```

Show the totals for the results for GHGs (kg CO2e), Jobs, and Value Added ($) by Region
```{r}
CO_result <- colSums(result$LCIA_d[grep("US-CO",row.names(result$LCIA_d)),])
RoUS_result <- colSums(result$LCIA_d[grep("RoUS",row.names(result$LCIA_d)),])
US_result <- CO_result+RoUS_result

kable(t(data.frame(CO=CO_result,RoUS=RoUS_result,`AllUS`=US_result)))
```
The full results can be viewed by sector and location
```{r}
kable(result$LCIA_d)
```
